# Use of Model to Predict Half-Lives for Novel PFAS

## Clear the workspace:
```{r clear_workspace}
rm(list=ls())
try(dev.off())
```

## Specify which RData files we are working with:
```{r configure_files}
MODELBUILDSUFF <- "JFW100322-noLogD" 
TRAININGSETSUFF <- "JFW012924"
WRITESUFF <- "JFW012924-noLogD" #This is the suffix for ongoing work. 
NEWPFASINFO <- "new_inv.csv"
NEWPFASOPERA <- "new_inv.csv"
NEWPFASENDO <- "EndoSubset-EPA_PFAS_DSSTox-Similarity_Tox21_102021_120723.csv"
SMILES.COL <- "QSAR_READY_SMILES"
```

## Configure Predictions
```{r randnumseed}
SPECIES <- "Human"
RNUMSEED <- "12345"
```

## Load necessary libraries
```{r load_libraries}
packages <- c("readxl","MLmetrics","stringr","scales","ggplot2")
# Note, the "character.only" argument is necessary here
sapply(packages, require,character.only=TRUE)
```

## Load list of new chemicals
Need "dtxsid", "name", "casrn", "formula", "mw"
```{r load_new_pfas}
setwd("c:/users/jwambaug/git/CompTox-PFASHalfLife")
pfastopred <- read.csv(paste0("NewPFAS/",
                              NEWPFASINFO))
dim(pfastopred)
# Require SMILES:
pfastopred <- subset(pfastopred, 
                     !is.na(pfastopred[,SMILES.COL]))
dim(pfastopred)
# Remove duplicated ID's:
pfastopred<- subset(pfastopred, !duplicated(DTXSID))
dim(pfastopred)

opera.vals <- read.csv(paste0("NewPFAS/",
                              NEWPFASOPERA))

endo.vals <- read.csv(paste0("NewPFAS/",
                              NEWPFASENDO))
```

# Load predictors:
```{r load_predictors}
setwd("c:/users/jwambaug/git/CompTox-PFASHalfLife")
load(file=paste0("RData/PFAS_QSAR_Predictors_Predictions_", 
                  MODELBUILDSUFF,
                  ".RData"))
predictors <- colnames(pfasdsredsc_ds)[4:(dim(pfasdsredsc_ds)[2]-2)]
cat(paste("Need the following descriptors:",
          paste(predictors, collapse=", ")
          ))
```

```{r add_endogenous_similarity}
needed.endo <- predictors[regexpr("TSPC",predictors)!=-1]
needed.endo <- gsub("TSPC_","",needed.endo)
needed.endo <- gsub("\\.","-",needed.endo)

endo.vals <- subset(endo.vals,ENDO.CASRN %in% needed.endo)
for (this.cas in needed.endo)
{
  for (this.id in unique(pfastopred$DTXSID))
    if (this.id %in% endo.vals$DTXSID)
    {
      col.name <- paste0("TPSC_",
                     gsub("-","\\.",this.cas))
      pfastopred[pfastopred$DTXSID==this.id, col.name] <-
      endo.vals[endo.vals$DTXSID==this.id &
                endo.vals$ENDO.CASRN==this.cas,
                "Tanimoto.Score..PubChem."]
  }
  pfastopred <- subset(pfastopred, 
                  !is.na(pfastopred[,col.name]))
}
```

```{r add_zero_for_serum_albumin}
pfastopred$Ka_perM_SerAlb_Han <- 0
```

```{r add_kidney_predictors}
setwd("c:/users/jwambaug/git/CompTox-PFASHalfLife")
kidney <- as.data.frame(read_excel(
  "Predictors/Oliver 1968_SupplementalData_JFW100322.xlsx", sheet =2))[1:11,]
kidney$Neph_Num <- as.numeric(kidney$Neph_Num)
kidney$BW <- as.numeric(kidney$BW)
kidney$KW <- as.numeric(kidney$KW)
kidney$GlomSA <- as.numeric(kidney$GlomSA)

kidney[!is.na(kidney$Neph_Num),"Log10Neph_Num"]=log10(kidney[!is.na(kidney$Neph_Num),"Neph_Num"])
kidney$Log10BW=log10(kidney$BW)
kidney$Log10KW=log10(kidney$KW)

if (!(tolower(SPECIES) %in% tolower(kidney$Mammal))) 
  stop("Kidney data for selected species unavailable")

this.row <- tolower(kidney$Mammal)==tolower(SPECIES)
```


 pfastopred.opera <- read.csv("NewPFAS/OPERA_desc.csv")
  
load(paste0("RData/PFAS_Training_Set_Endo_Disc_HLH_unscaled_",
            MODELBUILDSUFF,".RData", sep=""))
# Load pfasdsrd for name index:
pfasdsredHLH=pfasdsred
}
```


#Classification model predictions
```{r make_predictions}
load(file=paste0("RData/PFAS_QSAR_Predictors_Predictions_", 
                  MODELBUILDSUFF,
                  ".RData"))
set.seed(RNUMSEED)
pfaspred_HL <- predict(classmod4, newdata = pfastopred) 
```




#Lists of non-categorical variables that ended up in the training set; not categorical variables, including species, sex, and get added
AllVars <- names(pfasdsredHLH)

#2. Bring in DSSTox chemicals with Average Mass from dashboard
PFASdata <- subset(PFASdata, select=c("DTXSID", "PREFERRED_NAME", "CASRN"))

#3. Merge with Opera predictors
Opera <- read.csv(paste0("Predictors/DSSToxPFAS_DTXSID_SMILES_Oct21-1-smi_OPERA2.7Pred.csv"))

##Reduce to only variable columns of interest, and also pull AD info 
names(Opera)

#PUll out molecular weight and name the same as the training data
AVERAGE_MASS=Opera$MolWeight

Operapreds=names(Opera)[unlist(lapply(AllVars, function(x) {grep(x,names(Opera))} ))] #Find Opera columns
#[1] "LogP_pred"   "LogVP_pred"  "LogWS_pred"  "LogD74_pred"


#Can't pull AD's automatically because they aren't listed in a straight forwrd manner; just have to code in manualy
Operacols=c("MoleculeID","MolWeight", Operapreds, "AD_LogP", "AD_VP", "AD_WS","AD_KOA")

##Remove logp range column 
Operacols=Operacols[-c(which(Operacols=="LogP_predRange"))]

Opera=Opera[,c(which(names(Opera)=="RowID"),which(names(Opera)%in%Operacols))]

#Set name of average mass to same as model 
names(Opera)[which(names(Opera)=="MolWeight")]="AVERAGE_MASS"


##Merge with PFASdata
PFASdata1=merge(PFASdata, Opera, by.x="DTXSID", by.y="MoleculeID", all.x=TRUE)



####Check against training data
# Load pfads:
load(paste0("RData/PFAS_11Chemicals_QSARdataset_",trainingsetsuff,".RData", sep="")) #loads pfasds
load(paste0("RData/Training_Data_Reorder_For_Prediction_",modelbuildsuff,".RData")) #pfasdsorder
pfasds=pfasds[pfasdsreorder,]
head(pfasds)
#Select out predictor columns
idcols=which(names(pfasds)%in%c("DTXSID", "CASRN", "PREFERRED_NAME"))
pfasdsred1=data.frame(pfasds[,idcols], pfasdsred)




ps=PFASdata1[which(PFASdata1$DTXSID%in%pfasdsred1$DTXSID),]
pfasdsred1=pfasdsred1[pfasdsred1$DTXSID=="DTXSID3031862",]
ps=ps[ps$DTXSID=="DTXSID3031862",]
pfasdsred1
ps 

pfasdsred1=data.frame(pfasds[,idcols], pfasdsred)
ps=PFASdata1[which(PFASdata1$DTXSID%in%pfasdsred1$DTXSID),]
pfasdsred1=pfasdsred1[pfasdsred1$DTXSID=="DTXSID1037303",]
ps=ps[ps$DTXSID=="DTXSID1037303",]
pfasdsred1
ps 

pfasdsred1=data.frame(pfasds[,idcols], pfasdsred)
ps=PFASdata1[which(PFASdata1$DTXSID%in%pfasdsred1$DTXSID),]
pfasdsred1=pfasdsred1[pfasdsred1$DTXSID=="DTXSID7040150",]
ps=ps[ps$DTXSID=="DTXSID7040150",]
pfasdsred1
ps 

"DTXSID7040150"

#Good, have very close values for unscaled opera predictors and average mass to training set. 

#####
#Now, merge with similarity data from Prachi 
Endoinfo<-read.csv("Predictors/EndoSubset-EPA_PFAS_DSSTox-Similarity.csv")

# Extra chemicals for Richard Judson:
#Note thatin the file loaded into endo below, the second and third columns 
# (CASRN and DTXSID) refer to the endogenous compound,
#while both ENDO_CAS and ENDO_DTXSID give the DTXSID of the PFAS compound
endo <- read.csv("Predictors/EndoSubset-EPA_PFAS_DSSTox-Similarity_final_list_OPPT.csv")
#endo.melt <- melt(endo,id=c("CASRN","ENDO.DTXSID..EPA.PFAS."))
#endo.cast <- dcast(subset(endo.melt,variable=="Tanimoto.Score..PubChem."),ENDO.DTXSID..EPA.PFAS.~CASRN)
dim(endo)
endo <- subset(endo, !(ENDO.DTXSID..EPA.PFAS. %in% unique(Endoinfo$ENDO.DTXSID..EPA.PFAS.)))
dim(endo)

dim(Endoinfo)
Endoinfo <- rbind(Endoinfo,endo) 
dim(Endoinfo)
Endoinfo <- subset(Endoinfo,!duplicated(Endoinfo))
dim(Endoinfo)



###Grab the similarity values of the chemicals that are in the model; back calculate from abbreviations 
SimVarsPC=AllVars[which(grepl('TSPC', AllVars))]
####strip-off TS_ and replace "." with "-"
SimVarsPC=substring(SimVarsPC, 6)
SimVarsPC=str_replace_all(SimVarsPC, "[.]", "-")

###Next, pull similarity data from EndoinfoTotal1
Endosubset=Endoinfo[which(Endoinfo$CASRN%in%SimVarsPC),]
length(unique(Endosubset$ENDO.DTXSID..EPA.PFAS.)) #8384


###Fill in similarity values for each chemical in the model 
PFASdata2=PFASdata1
threshold=0.9                
for(i in SimVarsPC){
  sub=subset(Endosubset, CASRN==i, select=c("CASRN","ENDO.DTXSID..EPA.PFAS.", "Tanimoto.Score..PubChem."))
  names(sub)[2:3]=c("DTXSID", paste("TSPC_",i,sep=""))                         
  sub=sub[,-1]
  # Discretize similarity scores to 1/0:
  sub[,2]=ifelse(is.na(sub[,2]),NA,ifelse(sub[,2]>=threshold,1,0))
  PFASdata2=merge(PFASdata2, sub, all.x=TRUE,all.y=FALSE)
}
names(PFASdata2)
length(unique(PFASdata2$DTXSID))
length(which(is.na(PFASdata2$`TSPC_142-62-1`)==TRUE))

###Next add max similarity: Note, this would only be a max within the subset of available chemicals, so only 30 or so; this could be different if all chemicals were added
##This takes a bit of time because there are so many chemicals
#for(i in 1:length(PFASdata2$CASRN)){
#  a=subset(Endoinfo, ENDO.DTXSID..EPA.PFAS. == PFASdata2$DTXSID[i])
#  maxsim=max(a$Tanimoto.Score..PubChem.)
#  PFASdata2$MaxEndoPC[i]=maxsim  
#}

#PFASdata2
#simthreshold=0.9
#Endodisc=pfasds[, Endocols]
#range(apply(Endodisc, 2, FUN="max"))
#Endodisc=ifelse(Endodisc>=simthreshold, 1,0)

#Note that for some PFAS, there are no similar chemicals 
# could do this with an apply function too; for future reference
#sapply(PFASdata2$CASRN, function(x){
#  a=subset(Endoinfo, ENDO.DTXSID..EPA.PFAS. == x)
#  maxsim=max(a$Tanimoto.Score..PubChem.)
#  return(maxsim)
#})

###When there are no similar chemicals, this is set to zero. 
#PFASdata2$MaxEndoPC=ifelse(is.infinite(PFASdata2$MaxEndoPC), 0, PFASdata2$MaxEndoPC)

#########Check against training data
pfasdsred1=data.frame(pfasds[,idcols], pfasdsred)
ps=PFASdata2[which(PFASdata2$DTXSID%in%pfasdsred1$DTXSID),]
pfasdsred1=pfasdsred1[c(pfasdsred1$DTXSID=="DTXSID3031860" | pfasdsred1$DTXSID=="DTXSID3031864"),]
ps=ps[c(ps$DTXSID=="DTXSID3031860" | ps$DTXSID=="DTXSID3031864"),]
pfasdsred1[1:2,]
ps #Good, maintaining similar descretized similarity values  


#########
#Next, merge in kidney data to create slots for different species
###Adding in Sex to Kidney dataset 
load(paste0("RData/Kidney_Predictions_MultSpecies_",trainingsetsuff,".RData"))
load(paste0("RData/Kidney_Descriptors_in_Training_Set_",trainingsetsuff,".RData"))


kidney=dplyr::rename(kidney, "Species"="Mammal")
#"KW_BW_ratio"        "GlomTotSA_KW_ratio"
#"ProxTubLen"         "ProxTubVol"
kidney=subset(kidney,select=c("Species", "Type",kd))
kidney=kidney[-c(which(kidney$Species=="Whale" | kidney$Species=="Elephant")),]
#Parse down to the four parameters of interest
#Adding in categories for species, sex, and dosing approach here 
#Merging Kidney dataset into PFAS dataset
Sex=c("Male", "Female")
#Species=c("Human", "Monkey", "Rat", "Mouse")
DosingAdj=c("IV", "Oral", "Other")
Species=c("Human", "Monkey", "Rat", "Mouse", "Cattle", "Dog", "Rabbit", "Horse", "Chicken")
CatTab=expand.grid("Sex"=Sex, "Species"=Species, "DosingAdj"=DosingAdj)
kidney=merge(kidney, CatTab, by="Species", all=TRUE)

PFASdata3=PFASdata2
#Adding in Species and Sex to PFAS list
PFASdata3=merge(PFASdata3, Species)
PFASdata3=dplyr::rename(PFASdata3, "Species"="y")

PFASdata3=merge(PFASdata3, kidney, by="Species", all=TRUE)
dim(PFASdata3)
length(unique(PFASdata3$DTXSID))#8163 #9898



#########Check against training data
catcols=which(names(pfasds)%in%c("Species", "Sex", "Type", "DosingAdj"))
pfasdsred1=data.frame(pfasds[,idcols], pfasds[,catcols],pfasdsred)
ps=PFASdata3[which(PFASdata3$DTXSID%in%pfasdsred1$DTXSID),]
pfasdsred1=pfasdsred1[c(pfasdsred1$Species=="Rat" & pfasdsred1$Sex=="Male" & pfasdsred1$DosingAdj=="Other"),]
ps=ps[c(ps$Species=="Rat" & ps$Sex=="Male" & ps$DosingAdj=="Other"),]
ps
pfasdsred1[1:2,]

#Good, the rat and human data is the same  


#########

##Bring in data from Zhang)
#Z=read.xlsx("Predictors/Zhang et al.2013_PFAS_Protein_Dissociation Constants.xlsx")
#Z$Species="Human"
#Z=Z[-c(which(names(Z)=="ChemicalName"))]
#PFASdata3[which(PFASdata3$CASRN%in%Z$CASRN==FALSE),1:4] #chemicals in dataset not in Zhang
#dim(PFASdata3)
##De-salting some chemicals
##PFOS
##Note: Perfluoropentane sulfonic acid is not available
##    : Perfuoroheptansulfonic acid is not available
#fillin=Z[which(Z$ChemicalAbbreviation=="PFOS"),which(names(Z)=="CASRN")]
#Z[which(Z$CASRN==fillin),which(names(Z)=="CASRN")]="1763-23-1"
#fillin=Z[which(Z$ChemicalAbbreviation=="PFHxS"),which(names(Z)=="CASRN")]
#Z[which(Z$CASRN==fillin),which(names(Z)=="CASRN")]<-"355-46-4" 
#fillin=Z[which(Z$ChemicalAbbreviation=="PFBS"),which(names(Z)=="CASRN")]
#Z[which(Z$CASRN==fillin),which(names(Z)=="CASRN")]<-"375-73-5" 
##Z=subset(Z, select=c("CASRN", "Species", "Kd_hL_FABP"))
##Changing non-detects to 0
##Z$Kd_hL_FABP=as.numeric(ifelse(Z$Kd_hL_FABP=="ND", 0, Z$Kd_hL_FABP))
#
##PFASdata4=merge(PFASdata3, Z[,c("CASRN", "Kd_hL_FABP")],by="CASRN", all.x=TRUE, all.y=FALSE) #only merge by DTXSID; the data is only in humans here. 
#length(unique(PFASdata4$DTXSID))#9898

PFASdata4 <- PFASdata3
names(PFASdata4)

#Replace "-" with "."
names(PFASdata4)<-str_replace_all(names(PFASdata4), "-", ".") #change - to .

#Read in add in COC
COC=read.csv("Predictors/DSSTox_PFAS_ToxPrints_COC_aliphatic.csv") 
dim(COC)
ToxPrints <- read.csv("PFAS_Catalog/ToxPrints-COC.csv")
ToxPrints <- ToxPrints[,c("DTXSID","bond.COC_ether_aliphatic")]
dim(ToxPrints)
colnames(ToxPrints) <- colnames(COC)
COC <- rbind(COC,ToxPrints)
dim(COC)
COC <- subset(COC,!is.na(COC_aliphatic))
dim(COC)
COC <- subset(COC,!duplicated(DTXSID))
dim(COC)
save(COC, file="RData/PFAS_COC_DSSTox.RData")


PFASdata5=merge(PFASdata4, COC, by="DTXSID")

length(unique(PFASdata5$DTXSID))#6509 #8209

names(PFASdata5)

#Read in Data from Han
H=read_excel("Predictors/Han_et al.2012_AlbuminProteinBinding_Data.xlsx")
H$Reference_Ka_SerAlb="Han 2011"
H[which(H$PREFERRED_NAME=="Perfluorobutanesulfonate"), which(names(H)=="PREFERRED_NAME")]<-"Perfluorobutanesulfonic acid"

PFASdata6=merge(PFASdata5, H[,c("PREFERRED_NAME", "Species", "Ka_perM_SerAlb_Han")], by=c("PREFERRED_NAME", "Species"), all.x=TRUE, all.y=FALSE)
dim(PFASdata6)
length(unique(PFASdata6$DTXSID))#6509 #8297

####Check against training data
pfasdsred1=data.frame(pfasds[,idcols], pfasds[,catcols],pfasdsred)
ps=PFASdata6[which(PFASdata6$DTXSID%in%pfasdsred1$DTXSID),]
pfasdsred1=pfasdsred1[c(pfasdsred1$Species=="Human" & pfasdsred1$Sex=="Male" & pfasdsred1$DosingAdj=="Other" & pfasdsred1$DTXSID=="DTXSID3031862" ),]
ps=ps[c(ps$Species=="Human" & ps$Sex=="Male" & ps$DosingAdj=="Other" & ps$DTXSID=="DTXSID3031862" ),]
ps
pfasdsred1


######


AllVars%in%names(PFASdata6) #All continous variables are present 
IDcols=PFASdata6[,c("CASRN", "DTXSID", "Species")] #
OperaADcols=PFASdata6[,grep("AD", names(PFASdata6))]
length(unique(IDcols$DTXSID))#6509 #8297

PFASdata7=PFASdata6[,which(names(PFASdata6)%in%AllVars)] #only continous variables 
names(PFASdata7)



######
#Grab categorical sets and adjust the numerics to be on the same scale as the training data 
PFASdata7cat=subset(PFASdata6, select=c("Species","Sex",  "Type", "DosingAdj"))

#PFASdata6 is all numeric
sapply(PFASdata7, class)


######Fill in Average Values: will do at end for all numeric values 
#Find extent of missing data
missdat=apply(PFASdata7, 1, function(x){length(which(is.na(x)==TRUE))})
PFASdata7[343,]

hist(missdat)#Looks like the vast majority are missing at least 1 parameter,which is the FABP parameter. Probably should
#determine a better way to group those then. Instituting an adhoc decision to remove all chemicals in which >2 parameters
#have to be imputed via an average. 

removemissing=which(missdat>2) #Will need to remove from IDcols and categorical cols as well 
PFASdata7=PFASdata7[-removemissing,]

#Load averages of training set 
load(paste0("RData/Mean_SD_Trainingset_",modelbuildsuff,".RData"))

#Input averages of trainset for the rest for rest
for(i in 1:ncol(PFASdata7)){
  meancol=train_msd[1,which(names(train_msd)==names(PFASdata7)[i])]
  sub=which(is.na(PFASdata7[,i])==TRUE)  
  if(length(sub)>0){
    PFASdata7[sub,i]=meancol
  }}

which(is.na(PFASdata7)) #Should be zero 
#Make sure each cat is a factor 


  #Make sure categorical variables are factors,and remove chemicals with missing data
  PFASdata7cat$Sex<-as.factor(PFASdata7cat$Sex)
  PFASdata7cat$Type<-as.factor(PFASdata7cat$Type)
  PFASdata7cat$DosingAdj<-as.factor(PFASdata7cat$DosingAdj)
  PFASdata7cat=PFASdata7cat[-removemissing,]
  IDcols=IDcols[-removemissing,] 
  OperaADcols=OperaADcols[-removemissing,]
  length(unique(IDcols$DTXSID)) #6603; so this mainly results from the ad hoc decision to remove chemicals with >2 missing columns
  

  ###Scaling and centering on the training data mean and sd
  PFASdata7sc=PFASdata7
  for(i in 1:ncol(PFASdata7sc)){
    meancol=train_msd[1,which(names(train_msd)==names(PFASdata7sc)[i])]
    sdcol=train_msd[2,which(names(train_msd)==names(PFASdata7sc)[i])]
    PFASdata7sc[,i]=(PFASdata7sc[,i]-meancol)/sdcol  
  }


#Total Dataset to Predict
pfastopred=cbind(PFASdata7cat, PFASdata7sc)

#######Training data check
load(file=paste0("RData/PFAS_QSAR_Predictors_Predictions_", modelbuildsuff,".RData"))
pfasdsred1=data.frame(pfasds[,idcols], pfasds[,catcols],pfasdsredsc_ds)
ps=cbind(IDcols,PFASdata7cat,PFASdata7sc)
ps=ps[which(ps$DTXSID%in%pfasdsred1$DTXSID),]
pfasdsred1=pfasdsred1[c(pfasdsred1$Species=="Human" & pfasdsred1$Sex=="Male" & pfasdsred1$DosingAdj=="Other"),]
ps=ps[c(ps$Species=="Human" & ps$Sex=="Male" & ps$DosingAdj=="Other"),]
#Note; at least estimate for WS is very different; DTXSID3031860. So, I need to get new predictions
#from another source the model. 

set=NULL
for(i in unique(ps$DTXSID)){
  sub=ps[ps$DTXSID==i,which(names(ps)%in%c("DTXSID","TSPC_107.92.6",  "AVERAGE_MASS", "LogWS_pred", "LogP_pred", "LogVP_pred"))]
  sub2=pfasdsred1[pfasdsred1$DTXSID==i,which(names(pfasdsred1)%in%c("DTXSID","TSPC_107.92.6", "AVERAGE_MASS", "LogWS_pred", "LogP_pred", "LogVP_pred"))]
  order1=order(names(sub))
  sub=sub[,order1]
  order2=order(names(sub2))
  sub2=sub2[,order2]
  set=rbind(set,sub,sub2)
  }

set=rbind(ps[,c("DTXSID","TSPC_107.92.6", "AVERAGE_MASS", "LogWS_pred", "LogP_pred")],pfasdsred1[,c("DTXSID","TSPC_107.92.6", "AVERAGE_MASS", "LogWS_pred","LogP_pred")])
#set; very values for average mass and LogWS_pred


#Make model predictions 
#Regression model 
#load("RegressionModel_EndoSimDisc_HLH_DED061621.RData")
#set.seed(seed)
#pfastopred_regFull=predict(regmod, newdata = pfastopred)

#load("RegressionModel_PostRFE_HLH_DED061621.RData")
#set.seed(seed)
#pfastopred_regRed=predict(regmodred, newdata = pfastopred)

#Need to re-arrange the order


length(unique(IDcols$CASRN))
length(unique(IDcols$DTXSID))

#Making complete set for Tox21 
completedataset=data.frame(IDcols[,c("CASRN", "DTXSID")], OperaADcols, pfastopred,  "ClassPredFull"=pfastopred_classFull)

#Comaring RegPredFull with RegPredRed
#MAE(completedataset$RegPredFull, completedataset$RegPredRed)
#5442
#RMSE(completedataset$RegPredFull, completedataset$RegPredRed)
#7060 #pretty close agreement between the complete and reduced model 


save(completedataset, file=paste("RData/PFAS_HLpreds_HLH_HLHLSsc_DSSToxchemicals_coarse_imputation_",writesuff,".RData", sep="")) 
#load(paste("PFAS_HLpreds_HLH_HLHLSsc_DSSToxchemicals_coarse_imputation_",writesuff,".RData", sep=""))
length(unique(completedataset$CASRN))
# Load medians of the four bins:
load(paste("RData/Median_HLH_per_Bin_4_",
  modelbuildsuff,
  ".RData",sep=""))
agtabHLH
# Convert to elimination rate (1/h):
agtabHLH$kelim.h <- signif(log(2)/agtabHLH$HLH,3)
# Median Vd from data set is 0.202 L/kg BW
agtabHLH$CLtot.Lpkgbwpday <- signif(agtabHLH$kelim.h*0.202*24,3)
# PFOA health advisory CL 0.00014 L/kg BW/day #predicted by model here to be 0.000115
# PFOS health advisory CL 0.000081 L/kg BW/day
agtabHLH$CLtot.Lpkgbwpday
# 1 mg/kg BW/day dose rate -> Css (mg/L)
agtabHLH$Css.mgpL <- signif(1/agtabHLH$CLtot.Lpkgbwpday,3)

colnames(agtabHLH)[2] <- "HalfLife.h"
colnames(agtabHLH)[3] <- "kelim.1ph"

completedataset$HalfLife.h <- NA
completedataset$kelim.1ph <- NA
completedataset$CLtot.Lpkgbwpday <- NA
completedataset$Css.mgpL <- NA
for (this.bin in seq(1,4))
  for (this.col in c("HalfLife.h","kelim.1ph","CLtot.Lpkgbwpday","Css.mgpL"))
    completedataset[completedataset[,"ClassPredFull"]==this.bin, this.col] <-
      agtabHLH[agtabHLH[,"HLHBin4"]==this.bin, this.col]
      
  completedataset[completedataset$CASRN=="335-67-1" & completedataset$Species=="Human",]
  completedataset[completedataset$CASRN=="13252-13-6" & completedataset$Species=="Human",]
  

  

####Applicability Domain####
# This function implements the applicability domain estimation method of
# Roy et al. 2015 (https://doi.org/10.1016/j.chemolab.2015.04.013), 
# based on the descriptors that are scaled by those of the training set:
# "...After a descriptor column is standardized based on the corresponding mean 
# and standard deviation for the training set compounds only, if the 
# corresponding standardized value for descriptor i of compound k (Ski) is more 
# than 3, then the compound should be an X-outlier (if in the training set) or 
# outside AD (if in the test set) based on descriptor i." To be consistent with 
# Mansouri et al (https://doi.org/10.1186/s13321-018-0263-1) "1" indicates that
# a chemical is within the applicability domain and "0" indicates the chemical 
# is outside the domain. 
ADfunction=function(model, data)
{
  vars=names(model$trainingData)
  vars=vars[-c(length(vars))] #removes last column which is predictions
# Separate the numeric descriptors ("a") from the categorical descriptors ("b")
# and only analyze the numeric descriptors:
  cats=which(vars%in%c("Species", "Sex", "DosingAdj"))
  Vars1=vars
  if(length(cats>0)){
    Vars1=vars[-c(which(vars%in%c("Species", "Sex",  "DosingAdj")))]}
  a=data
  b=a[,c("CASRN", "DTXSID","Species", "Sex",  "DosingAdj")]
  a=a[,which(names(a)%in%Vars1)]
  a=abs(a)
# Calculate SI -- the standardized value for descriptor "I"
  SImin=apply(a,1,min) #Find the min, max, sd, and mean
  SImax=apply(a,1,max)
  SImean=apply(a,1,mean)
  SIsd=apply(a,1,sd)
  SIdf=data.frame(data,SImin, SImax, SImean, SIsd)
  SIdf$SI90=SIdf$SImean + (1.28 * SIdf$SIsd)
# Initialize domain variable, chemicals are assumed to be inside the AD 
# unless found not be: 
  SIdf$Domain=1 
  
# Use Roy et al. (2015) criteria for outside of domain of applicability:
  SIdf$Domain=ifelse(SIdf$SImin>3, 0, ifelse((SIdf$SImean+(1.28*SIsd))>3,0,1))
  
  return(SIdf)
}

CMFAD <- ADfunction(model=classmod4,data=completedataset)
#RMFAD <- ADfunction(model=regmod,data=completedataset)
#RMRAD <- ADfunction(model=regmodred,data=completedataset)

names(CMFAD)[which(names(CMFAD)=="Domain")]="ClassModDomain"
#names(RMFAD)[which(names(RMFAD)=="Domain")]="RegFullModDomain"
#names(RMRAD)[which(names(RMRAD)=="Domain")]="RegRedModDomain"

names(CMFAD)[which(names(CMFAD)=="SI90")]="ClassModSI90"
#names(RMFAD)[which(names(RMFAD)=="SI90")]="RegFullModSI90"
#names(RMRAD)[which(names(RMRAD)=="SI90")]="RegRedModSI90"

#DPFASwAD=data.frame(CMFAD[,-c(which(names(CMFAD)%in%c("SImin",  "SImax", "SImean","SIsd")))], RMFAD[,c("RegFullModSI90", "RegFullModDomain")], RMRAD[,c("RegRedModSI90", "RegRedModDomain")]) 
DPFASwAD=data.frame(CMFAD[,-c(which(names(CMFAD)%in%c("SImin",  "SImax", "SImean","SIsd")))]) 


DPFASwAD$AMAD <- apply(DPFASwAD[,Operacols[regexpr("AD",Operacols)>-1]],1,prod)

# Set reasonable sigfigs:
for (this.col in c(11:22,25:29))
{
  DPFASwAD[,this.col] <- signif(DPFASwAD[,this.col],4)
}

save(DPFASwAD, file=paste0("RData/Tox21_AllMods_ADindicate_", writesuff,".RData"))
write.csv(DPFASwAD, file=paste0("RData/Tox21_AllMods_ADindicate_", writesuff,".csv"))
#load(file=paste0("Tox21_AllMods_ADindicate_", writesuff,".RData"))
#DPFASwAD=read.csv(paste0("Tox21_AllMods_ADindicate_", writesuff,".csv"))
DPFASwAD[DPFASwAD$DTXSID=="DTXSID8047553" & DPFASwAD$Species=="Human",]

  DPFASwAD[DPFASwAD$CASRN=="335-67-1",]

#  PFOS=DTXSID3031865
#  PFOA=DTXSID40892486
#Gex=DTXSID70880215 

subset1=DPFASwAD[DPFASwAD$Sex=="Male" & DPFASwAD$DosingAdj=="Oral",]
subset1$Count=ifelse(is.na(subset1$Css.mgpL), NA,1)
aggregate(Count~Species, data=subset1, FUN="sum")  
  

#Note, only 1917 chemicals are listed as being in the domain of the reduced regresson model 
length(unique(DPFASwAD[DPFASwAD$ClassModDomain==1, "CASRN"])) #6117
#length(unique(DPFASwAD[DPFASwAD$RegFullModDomain==1, "CASRN"])) #3368
#length(unique(DPFASwAD[DPFASwAD$RegRedModDomain==1, "CASRN"])) #3368

###Domain tables
 dom=DPFASwAD
 dom=subset(dom, Species%in%c("Human","Monkey", "Rat", "Mouse", "Dog"))
 dom=subset(dom, Sex=="Female" & DosingAdj=="Other")
 CDtab=aggregate(ClassModDomain~DosingAdj+Sex + Species,data=dom, FUN = "sum")
# RRtab=aggregate(RegRedModDomain~Species,data=dom, FUN = "sum")
# RFtab=aggregate(RegFullModDomain~Species,data=dom, FUN = "sum")
 dom2=subset(dom, AD_LogP==1 & AD_VP==1 & AD_WS==1)
 CDtab2=aggregate(ClassModDomain~DosingAdj+Sex + Species,data=dom2, FUN = "sum")
# RRtab2=aggregate(RegRedModDomain~Species,data=dom2, FUN = "sum")
# 
# 
 domtab1=data.frame("Model"="Classification",CDtab,"All_Domains"=CDtab2[,"ClassModDomain"])
